---
title: "Stop and Frisk"
format: revealjs
editor: visual
---

# Introduction:

-   name: Nic Schwab

-   SDS department at Smith

-   Math department at Holyoke Community College

-   Develop a Data Science Program at HCC

-   email: nschwab\@smith.edu

# Why Data Science?

-   Math conference for teaching with Inquiry.

-   Math teachers will be the ones teaching data science.

-   Data Science lends itself to inquiry practices and active learning.

# Why R?

Lots of resources available to learn R.

see [Resources](sources/resources.qmd)

Many alternatives

-   [codap](https://codap.concord.org/)

-   [NuMPI](https://numpy.org/) with python

-   [google analytics](https://analytics.google.com/analytics/academy/course/6)

# Stop, Question, Frisk

-   Started in the 1990s

-   New York City Police Practice

    -   Police only need reasonable suspicion.

-   Many articles about the targeting of Black and Latino young men.

-   Studies have shown that this has led to mistrust in the police.

# More about about me

I'm not Black.

I'm am Hispanic.

I'm White

I am not from New York City.

# Race Exists

Let's acknowledge that Race exists and plays a part in the way human being interact with each other.

# Schedule:

::: incremental
1.  Introduce you to the data.

2.  Pose a question that I have about stop and frisk.

3.  Give you Data Frames with data to answer the question one way.

    i.  You'll order them with a partner.

    ii. You'll make up puesdo code to explain data transformations.

    iii. I'll give you the real code.

4.  We'll repeat the process of data wrangling in R.
:::

# Questions so far?

# The Data

[Here's](https://www.nyc.gov/site/nypd/stats/reports-analysis/stopfrisk.page) NYC's website with the stop and frisk data.

Let's consider 2011 data

::: nonincremental
-   Each column is a variable

-   and each row is a stop
:::

Data Dictionaries exist for older data.

New Data variables are more clear.

The variable names are inconsistent across years.

# SQF History in a graph

```{r stop-frisk-years}
#| echo: TRUE
#| warning: FALSE
#| include: FALSE

library(tidyverse)
library(readxl)
options(scipen = 10)



```

```{r }
#| eval: False

# Read in each csv file. This is a lot 

# With purr csv

file_paths_csv <- fs::dir_ls("../data/2003_2015_csv_files")

file_contents_csv<- file_paths_csv |>
  map(function (path){
    read_csv(path)
  })

# with purr excel

file_paths_xl <- fs::dir_ls("../data/2016_2022_xl_files")

file_contents_xl<- file_paths_xl |>
  map(function (path){
    read_excel(path)
  })

# https://www.youtube.com/watch?v=An1bUIg-nVM

#See the number of rows i.e stops in each year
pre_2016 <- sapply(file_contents_csv, nrow)
post_2016 <- sapply(file_contents_xl, nrow)


```

```{r graph-of-stop}

years <- c(2003, 2004, 2005, 2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020, 2021, 2022 )
stops_frisks <- c(160851, 313523, 398191, 506491, 472096, 540302, 581168,  601285, 685724, 532911,  191851, 45787, 22563, 12405, 11629,  11008,   13459,9544 , 8947,  15102)  

sqf <- data.frame(years,stops_frisks)



sqf |> 
  ggplot(aes(years,stops_frisks)) +
  
  geom_rect(aes(xmin=2002,xmax=2014,ymin=0,ymax=Inf),fill="#003585",alpha=0.01)+
  geom_rect(aes(xmin=2014,xmax=2021,ymin=0,ymax=Inf),fill="#FFFFFF",alpha=0.01)+
  geom_rect(aes(xmin=2021,xmax=2022,ymin=0,ymax=Inf),fill="#FF6600",alpha=0.01)+
  geom_line()+
  xlab("Year")+
  ylab("Citizens Stopped ('000s)")+
  labs(title="")+
  theme_minimal()+
  scale_y_continuous(breaks = c(0,350000,700000), labels = c("0","350","700"))+
  scale_x_continuous(breaks = c(2003,2013,2022))+
  geom_text(label="Bloomberg", x = 2004.5, y = 40000, color="gray35" ) +
  geom_text(label="de Blasio", x = 2016, y = 40000, color="gray35" ) +
  geom_text(label="Adams", x = 2022, y = 40000, color="gray35" ) +
  geom_text(label=". NYCLU \n suit", x = 2013, y = 650000, color="gray35" ) +
  geom_vline(xintercept=2012, linetype = 3 )
  
```

# Questions

> ::: incremental
> 1.  Which racial groups were most likely to be stopped by police officers 2011?
>
> 2.  What percent of the stops resulted in an arrest in 2011?
>
> 3.  How do the answers to these questions change in 2021?
> :::

# Here are your data frames

1.  Little envelopes Data 2011

-   Put the Data Frames in Order.

-   Be sure you can explain the process from one DF to another.

2.  Little envelopes Functions

-   After the frames are in Order put the code arrows between each frame.

