---
title: "Proportion of stops 2011"
author: "Schwab"
format: pdf
editor: visual
execute: 
  echo: false
---

## Envelope 2011

```{r libraries-and-data}
#| include: FALSE
#| echo: FALSE

library(tidyverse)
library(readr)
library(readxl)
library(knitr)
library(kableExtra)

library(tidyverse)

sqf_url <- "https://www1.nyc.gov/assets/nypd/downloads/zip/analysis_and_planning/stop-question-frisk/sqf-2011-csv.zip"
temp <- tempfile()
download.file(sqf_url, temp)
sqf_zip <- unzip(temp, "2011.csv")
sqf_2011 <- read.csv(sqf_zip, stringsAsFactors = FALSE) 
sqf_2011_race_cat <- read.csv("https://raw.githubusercontent.com/lindsaypoirier/STS-101/master/Data/SQF/sqf_race_categories.csv", stringsAsFactors = FALSE) 
rm(sqf_url)
rm(temp)
rm(sqf_zip)

```

# Guide

These materials were made for the NECOMMIT 2023 conference keynote.

This document is meant to be cut up and put into envelopes to distribute to participants. One envelope per group.

This guiding questions is :

> 1.  Which racial groups were most likely to be stopped by police officers 2011?
>
> 2.  What percent of the stops resulted in an arrest in 2011?
>
> 3.  How do the answers to these questions change in 2021?

# Start 2011

\newpage

```{r}
total_stops <-nrow(sqf_2011)
```

Basic Data Frame

```{r}
sqf_2011 |>
  select(year,sex,race,arstmade,crimsusp,sumissue,frisked,searched,pistol)|>
  head(15) |>
  kable(format="latex", booktabs = TRUE) |>
  kable_styling(latex_options = "striped")
```

\newpage

```{r select}
sqf_2011 <- sqf_2011  |>
  select(sex,race, arstmade 
         )

sqf_2011 |> 
  head(20) |>
  kable(format="latex", booktabs = TRUE) |>
  kable_styling(latex_options = "striped")
```

\newpage

```{r recode_race}

sqf_2011$race<-
  recode(
    sqf_2011$race, 
    "A" =	"API",
    "B"	= "Black",
    "I"=	"Am Indian",
    "P"=	"Hispanic",
    "Q"=	"Hispanic",
    "W"=	"White",
    "X"=	"Unknown",
    "Z"=	"Other",
    .default = NA_character_
  )

sqf_2011 |> 
  head(20) |>
 kable(format="latex", booktabs = TRUE) |>
  kable_styling(latex_options = "striped")
```

\newpage

```{r recode_arstmade}
# There is no missing data for arrestmade. 

sqf_2011$arstmade<-
  recode(
    sqf_2011$arstmade, 
    "Y" =	1,
    "N"	= 0
  )

sqf_2011 |> 
  head(20) |>
 kable(format="latex", booktabs = TRUE) |>
  kable_styling(latex_options = "striped")
```

\newpage

```{r count }
# This removed NAs I should leave this out of the conference.  

# sqf_2011 |> 
#   filter(!is.na(race) & race != "Other") |>
#   rename(race_na_removed = race)
# 
# sqf_2011  |>
#   kable(format="latex", booktabs = TRUE) |>
#   kable_styling(latex_options = "striped")
```

\newpage

```{r}
#count the total stops
sqf_2011 |>
  group_by(race) |>
  summarise(stops_by_race = n() ) |>
  

  kable(format="latex", booktabs = TRUE) |>
  kable_styling(latex_options = "striped")
```

\newpage

```{r percent_stopped}

#This code finds the percent stopped

sqf_2011 |> 
  group_by(race) |>
  summarise(stops_by_race = n(),
            percent_stopped = n() / total_stops *100
  )|>
  
  kable(format="latex", booktabs = TRUE) |>
  kable_styling(latex_options = "striped")
```

\newpage

```{r percent_arrested_of_stopped}
#This code finds of the percent stopped how many are arrested. 

sqf_2011 |> 
  group_by(race) |>
  summarise(stops_by_race = n(),
            percent_stopped = n() / total_stops *100,
            percent_arrested = sum(arstmade)/n() *100
  )|>
  
  kable(format="latex", booktabs = TRUE) |>
  kable_styling(latex_options = "striped")
```

\newpage

```{r arrange}
#This code arranges by percent stopped. 

sqf_2011 |> 
  group_by(race) |>
  summarise(stops_by_race = n(),
            percent_stopped = n() / total_stops *100,
            percent_arrested = sum(arstmade)/n() *100
  )|>
  arrange(percent_arrested) |>
  
  kable(format="latex", booktabs = TRUE) |>
  kable_styling(latex_options = "striped")
```

```{r graph}
sqf_2011 |> 
  group_by(race) |>
  summarise(stops_by_race = n(),
            percent_stopped = n() / total_stops *100,
            percent_arrested = sum(arstmade)/n() *100
  )|>
  arrange(percent_arrested) |>
  ggplot(aes(x=race,y=percent_stopped)) +
  geom_col(fill="red")+
  theme_minimal()

```

```{r graph2}
# this graph is misleading. 

# sqf_2011 |> 
#   group_by(race) |>
#   summarise(stops_by_race = n(),
#             percent_stopped = n() / total_stops *100,
#             percent_arrested = sum(arstmade)/n() *100
#   )|>
#   arrange(percent_arrested) |>
#   ggplot()+
#   geom_col(aes(x=race,y=percent_arrested),   fill="blue", width = 0.25, position = position_nudge(x=0.15))+
#   geom_col(aes(x=race,y=percent_stopped), fill="red",width = 0.25, position = position_nudge(x=-0.15))+
#   labs(
#         y="", 
#        title = "Percent Stopped and Percent Arrested",
#        subtitle = "NYC SQF 2011 Data",
#        caption = "The red bars show percent of people stopped of the total.\n The blue bars show the percent arrested by stop.")+
#   theme_minimal()

```
